Curve counting in genus one: elliptic singularities and the splitting formula

### 2019-06-26 Reduced Gromov-Witten Theory of Hypersurfaces in Genus One.tex

\title{Absolute, relative and rubber Gromov--Witten theory of hypersurfaces in genus one}
\author{Luca Battistella, Navid Nabijou and Dhruv Ranganathan}
\date{\thismonthyear}

@@

@@

\subsubsection{From $\widetilde\Dcal$ to $\Dcal$}

\subsubsection{Factorisation} Assume $m\geq 2$. Over $\mathcal P$ we have a tautological sequence:
\[\OO_{F}(-1)\hookrightarrow p^*(F)\xrightarrow{+\operatorname{d}f}\operatorname{ev}_E^*(TH),\]
where $p\colon\mathcal P=\PP(F)\to\widetilde\Ecal$, and $+\operatorname{d}f$ should be interpreted as follows: $k$ copies of $\mathbb T_R$ are taken to account - via the preferred isomorphisms - for all factors of the first product in \eqref{eq:product}; then, the differential of the universal map to $\PP^N$ is applied to each tangent vector (possibly including some of the contracted rational bubbles in the boundary), and the sum of these contributions is taken in $T_{f(E)}\PP^N$ - it actually lies in $T_{f(E)}H$ because all the normal contributions are trivial by assumption. The composition is a section $s$ of $G=\OO_F(1)\otimes \operatorname{ev}_E^*(TH)$, and it vanishes precisely where $f$ factors through the Smyth's singularity determined by the point in the fiber of $p$. Generically on $\widetilde\Ecal$, this is exactly what we are after.

@@

\subsection{Tautological contributions} Expressing the pushforward of the Chern classes of $\OO_F(1)$ and $\OO(\widetilde\Delta_i)$ in terms of tautological classes on the fiber product underlying $\Ecal$, get:
\begin{cor}
 After pushforward to $\Ecal$, the contribution of $\Dcal$ to the invariants can be computed in terms of tautological integrals on moduli spaces of maps with lower numerics.
\end{cor}

>>>>

\title{Curve counting in genus one: elliptic singularities and the splitting formula}
\author{Luca Battistella, Navid Nabijou and Dhruv Ranganathan}
\date{\thismonthyear}

@@

\begin{lemma}\label{lem:m=1}
 If the maximal contact order at the core is $m=1$, the corresponding combinatorial type does not contribute to the invariants.
\end{lemma}
\begin{proof}
 It follows once again from a dimensional argument.
\end{proof}

@@

\subsubsection{From $\widetilde\Dcal$ to $\Dcal$} Let $F=\bigoplus_{v\in \{V_{0,\delta}\}\cup V_{>0,\delta}}\OO(\delta_v)$. On $\mathcal P$ we have a tautological sequence:
\[\OO_{F}(-1)\hookrightarrow p^*(F)\xrightarrow{+\operatorname{d}f}\operatorname{ev}_E^*(TH),\]
where $p\colon\mathcal P=\PP(F)\to\widetilde\Ecal$, and $+\operatorname{d}f$ should be interpreted as follows: for every $v\in \{V_{0,\delta}\}\cup V_{>0,\delta}$, as many copies of $\OO(\delta_v)$ are considered as there are components at distance $\delta$ from the core along the path $v$, and they are identified according to the preferred isomorphisms; then, the differential of the universal map to $\PP^N$ is applied to each tangent vector (possibly including some contracted rational bubbles on the path), and the sum of these contributions is taken in $T_{f(E)}\PP^N$ - it actually lies in $T_{f(E)}H$ because all the normal contributions can be assumed to be trivial by Lemma \ref{lem:m=1}. The composite, regarded as a section $s$ of $G=\OO_F(1)\otimes \operatorname{ev}_E^*(TH)$, vanishes precisely where $f$ factors through the Smyth's singularity determined by the point in the fiber of $p$. Generically on $\widetilde\Ecal$, this is exactly what we are after. On the boundary, though, it may happen that a simultaneous degeneration in the base and the fiber makes this factorisation trivial, in the sense that $\bar f$ would be constant on all the branches of the elliptic singularity.

\begin{lemma}
 For every $V_B\subseteq V_\delta$ and $V_F=V_\delta\setminus V_B$, $s$ vanishes along the loci where the factors corresponding to $V_B$ in $\widetilde \Ecal$ degenerate so that $f_v\colon R_v\to\PP^N$ is constant in a neighbourhood of $q_v$, and the fiber coordinates $(\lambda_{v^\prime})_{v^\prime\in V_F}$ along $p\colon\mathcal P\to\widetilde \Ecal$ are set to be $0$.
\end{lemma}
Call these loci $\Xi_i,i\in I$ (compare with \cite[\S 3.2]{VZ} for an analogous phenomenon in the non-relative case); it is clear that they may be in excess with respect to the expected codimension of $V(s)$, that is $N-1$. These loci are logarithmic substrata; therefore, there exists a logarithmic blow-up that principalizes them. We claim that $\widetilde\Dcal$ is (possibly a refinement of) such a blow-up. 
\begin{lemma}
 The pullback of $\Xi_i$ is principal in $\widetilde\Dcal$.
\end{lemma}
\begin{proof}
 Let us analyse the ideal of $\Xi_i$: it is generated by
 \begin{enumerate}[leftmargin=.5 cm]
  \item equations on $\widetilde\Ecal$ cutting the locus where $R_v$ degenerates so that $f_v$ is constant on a neighbourhood of $q_v$, $v\in V_B$: these correspond to the tropical functions $\delta_v-\operatorname{dist}(q_v,\circ)$;
  \item equations on the fiber of $p$ for the linear subspace corresponding to $v\in V_F$: tropically, these correspond to the $\ell_v$ discussed above.
 \end{enumerate}
By construction, on $\widetilde \Dcal$ it is always possible to tell which of these tropical lengths is the shortest (the circle we have drawn on $\Xi$ is degenerate - it passes through no vertex of positive horizontal degree -, therefore we need to decide for a strictly larger circle in order to lift to $\widetilde D$), which corresponds to principalising the ideal (see Figure \ref{fig:principalisation}).
\end{proof}

\begin{figure}
\begin{tikzpicture}[ipe import]
  %\useasboundingbox (0, 0) rectangle (595, 842);
  \draw
    (64, 720)
     -- (112, 736)
     -- (96, 752);
  \draw[->]
    (64, 704)
     -- (128, 704);
  \pic
     at (64, 704) {ipe disk};
  \pic
     at (64, 720) {ipe disk};
  \pic
     at (96, 752) {ipe disk};
  \draw[red, dotted]
    (60.6969, 708.808)
     arc[start angle=-161.5651, end angle=-71.5651, x radius=53.2078, y radius=-65.2105];
  \pic[fill=white]
     at (112, 736) {ipe fdisk};
  \draw
    (192, 720)
     -- (240, 736)
     -- (224, 752);
  \draw[->]
    (192, 704)
     -- (256, 704);
  \pic
     at (192, 704) {ipe disk};
  \pic
     at (192, 720) {ipe disk};
  \pic
     at (224, 752) {ipe disk};
  \pic[fill=white]
     at (240, 736) {ipe fdisk};
  \draw
    (112, 736)
     -- (128, 736);
  \draw
    (240, 736)
     -- (256, 736);
  \draw
    (208.6374, 743.6181)
     -- (224.4352, 750.8588);
  \draw
    (208.6374, 758.7577)
     -- (225.0934, 751.517);
  \pic
     at (207.9791, 742.9599) {ipe disk};
  \pic
     at (207.9791, 758.7577) {ipe disk};
  \pic
     at (208, 704) {ipe cross};
  \pic
     at (224, 704) {ipe cross};
  \pic
     at (240, 704) {ipe cross};
  \pic
     at (96, 704) {ipe cross};
  \pic
     at (112, 704) {ipe cross};
  \draw[red, dotted]
    (231.0126, 755.9018)
     { [rotate=-17.595] arc[start angle=135, end angle=225, x radius=21.7093, y radius=22.6274] };
  \draw[red]
    (188.9249, 712.0225)
     arc[start angle=-156.5092, end angle=-115.3692, x radius=85.088, y radius=-96.462];
\end{tikzpicture}
\caption{On the left, a generic point of $\mathcal P$. On the right, a point of $\Xi$. The dotted circle is the one we draw on $\mathcal P$. The continuous circle represents one of the possible choices of $\delta$ enlarging the dotted one (a stratum of $\widetilde\Dcal$).}\label{fig:principalisation}
\end{figure}

Finally, on $\widetilde\Dcal$ (or, in fact, on an intermediate blow-up $\widetilde{\mathcal P}\to\mathcal P$) there are exceptional divisors $\widetilde \Xi_i$ corresponding to $\Xi_i$ in $\mathcal P$. Since the $\Xi_i$ exhaust all the excess-dimensional components of the vanishing locus of $s\in\Gamma(\mathcal P,G)$, the latter induces a section $\tilde s\in\Gamma(\widetilde\Dcal,G(-\sum_i\widetilde\Xi_i))$ by pullback and twisting, whose vanishing locus is dimensionally transverse to the boundary, and whose open part coincides with the generic point of $\Dcal$, therefore $V(\tilde s)=\Dcal$.

Notice that, expressing the pushforward of the Chern classes of $\OO_F(1)$ and $\OO(\widetilde\Xi_i)$ in terms of tautological classes on the fiber product underlying $\Ecal$, we get:
\begin{cor}
 After pushforward to $\Ecal$, the contribution of $\Dcal$ to the invariants can be computed in terms of tautological integrals on moduli spaces of maps with lower numerics.
\end{cor}
\begin{comment}
\subsubsection{Factorisation} Assume $m\geq 2$. Over $\mathcal P$ we have a tautological sequence:
\[\OO_{F}(-1)\hookrightarrow p^*(F)\xrightarrow{+\operatorname{d}f}\operatorname{ev}_E^*(TH),\]
where $p\colon\mathcal P=\PP(F)\to\widetilde\Ecal$, and $+\operatorname{d}f$ should be interpreted as follows: $k$ copies of $\mathbb T_R$ are taken to account - via the preferred isomorphisms - for all factors of the first product in \eqref{eq:product}; then, the differential of the universal map to $\PP^N$ is applied to each tangent vector (possibly including some of the contracted rational bubbles in the boundary), and the sum of these contributions is taken in $T_{f(E)}\PP^N$ - it actually lies in $T_{f(E)}H$ because all the normal contributions are trivial by assumption. The composition is a section $s$ of $G=\OO_F(1)\otimes \operatorname{ev}_E^*(TH)$, and it vanishes precisely where $f$ factors through the Smyth's singularity determined by the point in the fiber of $p$. Generically on $\widetilde\Ecal$, this is exactly what we are after.

@@

\subsection{Tautological contributions} Expressing the pushforward of the Chern classes of $\OO_F(1)$ and $\OO(\widetilde\Delta_i)$ in terms of tautological classes on the fiber product underlying $\Ecal$, get:
\begin{cor}
 After pushforward to $\Ecal$, the contribution of $\Dcal$ to the invariants can be computed in terms of tautological integrals on moduli spaces of maps with lower numerics.
\end{cor}
\end{comment}
